### 2016-02-12 prepReport/intro.tex
Implemented the Ivo suggested corrections

\chapter{Introduction}\label{chap:intro}

\section{Background}

\section{Objectives}

\section{Motivation}

>>>>

\chapter{Introduction}\label{chap:intro}
Brief overview of the report:

In Chapter~\ref{chap:intro}, I will present a conceptual background to the problem at hand, exposing the background that surrounds it~\ref{sec:intro_back}, the motivation for the desired solution~\ref{sec:intro_mot}, its objectives~\ref{sec:intro_obj} and, lastly, the people that will help me achieve them~\ref{sec:intro_people}.

In Chapter~\ref{chap:theorBack}, the theoretical foundations will be layed out. Section~\ref{sec:tb_basicCon} presents the basic concepts of Machine Learning, and in the following sections, the theoretical details of ANNs, RNNs and LSTMs. Section~\ref{sec:training_lstm} provides a quick explanation of the training algorithm that will be used in the final solution, which will be outlined in Section~\ref{sec:propSol}.

In Chapter~\ref{chap:stateArt}, the state of the art of LSTMs, their applications~\ref{sec:sa_apps}, their hardware implementations~\ref{sec:sa_hardware} and the current work regarding the chosen training algorithm~\ref{sec:sa_training} are presented.

In Chapter~\ref{chap:workPlan}, the work plan for the Dissertation work is detailed as \emph{tasks} in Section~\ref{sec:wp_tasks_td}, and their temporal arrangement is pictured in a Gantt chart in Section~\ref{sec:wp_tasks_gantt}. Furthermore, the description of the Hardware and Software resources that I will be utilizing during the course of my work is done in Section~\ref{sec:wp_sh}

Finally, in Chapter~\ref{chap:earlyWork} I will present some of the work that I have already concluded.

\section{Background}\label{sec:intro_back}
Artificial Neural Networks are one of the most popular models in the field of Machine Learning. As the name suggests, their operation is inspired by the operation of the building blocks of our brain, the \textbf{neurons}. In spite of its high performance, one of their shortcomings is the fact that they cannot retain temporal dependences among incoming data samples, thus not being suitable to process time-series data, such as audio, video or other kinds of time-varying data streams, where current inputs have a high temporal dependence with previous and future inputs.

To address this issue, several algorithms have been used, such as Hidden Markov Models (HMM) or Recurrent Neural Networks (RNN), but both of these methods fail to recall temporal dependences that extend over an large period of time, for the reasons that we will understand in Sections~\ref{sec:theorBack_rnn} and~\ref{sec:theorBack_lstm}.Thus, in 1997 Hochreiter et al. proposed a novel RNN structure~\cite{Hoch97}, the Long Short-Term Network (LSTM), where a memory cell was introduced, and the input/output/read/write access is controlled by individual gates that are activated both by the incoming data samples, but also by the outputs from the previous time-step (it is an RNN after all). They are one of the state of the art methods in Deep Learning nowadays, as we can attest in Section~\ref{sec:sa_apps}.

\section{Motivation}\label{sec:intro_mot}
Hitherto, and to the best of my knowledge, all of the published applications that use LSTM are software based, but the parallel nature of the structure hollers for a dedicated hardware realization that can dramatically increase its performance, something that  has only recently been done once~\cite{Chang15} (see Section~\ref{sec:sa_hardware} for further details), and although it improves the processing time when compared to a naïve software solution, it still lacks the ability to perform on-chip learning, and the learning process is performed offline, in a normal CPU, when it also could be speeded up in a dedicated hardware structure.

All these techniques are generally implemented in mainstream processors, making use of general high-level or low-level languages, where all the real parallelism is limited to the number of simultaneous threads that we can run on each physical core, which up to the date generally have between 2-8 cores (mobile devices and general use personal computers),

In order to parallelize the computations to the fullest extent, a solution is to port it both to a Graphics Processing Unit (GPU) or even a Field-Programmable Gate Array (FPGA), but the portin process is not entierely automatic and to have the least performance drop possible, it has to be explicitly programmed in CUDA/OpenCL for GPUs, and a Hardware Description Language for FPGAs, with an increasing level of complexity and low-level details. This way, it is necessary to provide frameworks that can allow an FPGA to quickly reconfigure itself to run these kind of networks on demand, for a particular task that requires them, achieving a supposedly lower computation time, and unburdening the CPU from running it, thus saving performance for running other tasks related with the Operating System, for instance. Furthermore, when processing an incoming stream of highly dimensional data, or with high throughput, a CPU solution might not be scalable, and could benefit greatly from a dedicated hardware implementation.

\section{Objectives}\label{sec:intro_obj}
Taking into account the considerations presented in~\ref{sec:intro_mot}, I propose to develop a hardware implementation of an LSTM Network, \emph{with on-chip learning}, improving the performance, capability and flexibility of the existing solution of~\cite{Chang15}. Moreover, I also propose to benchmark this solution and try to compare it with the software performance results of~\cite{Tavear13} and, as a secondary objective, try to use my developed structure to replicate the some of the applications portrayed in~\ref{sec:sa_apps}. Lastly, I will work to make this structure reconfigurable on the go, and trying to minimize the reconfiguration time.

\section{People Involved}\label{sec:intro_people}
Besides me, the candidate to the Master's Degree, there are two more people involved, namely

\begin{itemize}
    \item \textbf{Supervisor} -- The \href{https://sigarra.up.pt/feup/pt/func\_geral.formview?p\_codigo=210963}{Professor João Canas Ferreira}, auxiliary professor at the Faculty of Engineering of the University of Porto.
    \item \textbf{Second Supervisor} -- \href{http://www.cl.cam.ac.uk/~ijpdmt2/}{Ivo Timóteo}, a Computer Science PhD candidate at Cambridge University, UK, in the field of Artificial Inteligence.
\end{itemize}

### 2016-02-12 prepReport/stateArt.tex
Implemented the Ivo suggested corrections

    \item \textbf{Biomedical Applications} -- this network architecture was used in a protein homology detection scheme~\cite{Hochreiter07} using the SCOP 1.53 benchmark dataset, displaying a staggering performance when compared to other methods. Similarly, a recent article from 2015 by Sønderby et al. suggested the use of standalone LSTM and also Convolutional LSTM to perform subcellular localization of proteins, given solely their protein sequence, with an astounding accuracy of 0.902~\cite{Sonderby15}

@@

    \item \textbf{Video and Image Analysis} -- Vinyals et al., at Google, proposed and LSTM network for \textbf{image captioning}, preceded by a Convolutional Neural Network to apply preprocessing to the images~\cite{Vinyals14}. This way, the LSTM works as \emph{sentence generator}, that captions the images, with state of the art performance. 
        Besides image captioning, video captioning is also an interesting topic. Venugopalan et al. recently proposed a CNN + LSTM architecture to translate video sequences to natural language~\cite{Donahue14}, using the Microsoft Research Video Description Corpus as a dataset.Tthere are other similar studies that combine image and video captioning~\cite{Donahue14_2}.

@@

\section{Final Overview}\label{sec:overview}
As we could attest from this small literature survey, although there is already an hardware implementation os LSTM, there is still a good deal of room for improvement by adding on chip-learning to the system, and also to restric it to a smaller use of the FPGA resources, allowing it to accomodate a more complex network. Furthermore,~\cite{Tavear13} shows that, at least for that particular case, the LSTM network doesn't suffer a great performance impact from using SPSA training, as opposed to the more common BPTT, and~\cite{Maeda05} showed that an SPSA hardware implementation is feasible. These three conclusions combined indicate that the idea of a hardware LSTM network with on-chip learning using SPSA is possible.

>>>>

    \item \textbf{Biomedical Applications} -- this network architecture was used in a protein homology detection scheme~\cite{Hochreiter07} using the SCOP 1.53 benchmark dataset, displaying a staggering performance when compared to other methods. Similarly, a recent article from 2015 by Sønderby et al. suggested the use of standalone LSTM and also Convolutional LSTM to perform sub cellular localization of proteins, given solely their protein sequence, with an astounding accuracy of 0.902~\cite{Sonderby15}

@@

    \item \textbf{Video and Image Analysis} -- Vinyals et al., at Google, proposed and LSTM network for \textbf{image captioning}, preceded by a Convolutional Neural Network to apply preprocessing to the images~\cite{Vinyals14}. This way, the LSTM works as \emph{sentence generator}, that captions the images, with state of the art performance. 
        Besides image captioning, video captioning is also an interesting topic. Venugopalan et al. recently proposed a CNN + LSTM architecture to translate video sequences to natural language~\cite{Donahue14}, using the Microsoft Research Video Description Corpus as a dataset.There are other similar studies that combine image and video captioning~\cite{Donahue14_2}.

@@

\section{Final Overview}\label{sec:overview}
As we could attest from this small literature survey, although there is already an hardware implementation os LSTM, there is still a good deal of room for improvement by adding on chip-learning to the system, and also to restrict it to a smaller use of the FPGA resources, allowing it to accomodate a more complex network. Furthermore,~\cite{Tavear13} shows that, at least for that particular case, the LSTM network doesn't suffer a great performance impact from using SPSA training, as opposed to the more common BPTT, and~\cite{Maeda05} showed that an SPSA hardware implementation is feasible. These three conclusions combined indicate that the idea of a hardware LSTM network with on-chip learning using SPSA is possible.

### 2016-02-12 prepReport/theorBack.tex
Added a recurrent neuron figure
Implemented the Ivo suggested corrections

Let us consider a practical example. For instance, suppose we want to build a program that given an input audio waveform representation of a spoken word, it matches it into a particular word of a dictionary. We could, of course, devise a set of rules and exceptions for each word analysing some of its features (perhaps the Fourier representation of each one, and, from it, manually finding the appropriate rules for each), but apart from it being a very complex task, it wouldn't be a scalable solution, given the enormous number of words in each language. The approach taken by Machine Learning is diametrically different, and instead of manually processing each waveform, we build a large dataset, of size $N$, containing the waveforms of several words $\left[ \mb{x}_1 \; \mb{x}_2 \;\mb{x}_2 \; \dots \mb{x}_N \right]$ -- we call this dataset the \textbf{training dataset} -- and we feed it to our model. Each of the $i$-th data point was previously labelled, and in fact we feed each training data point $\mb{x}_i$ \emph{along} with its corresponding label $t_i$, so that the model can adapt its parameters accordingly to the \emph{target value} it is supposed to classify. This set of labels $\mb{t} = \left[ t_1 \; t_2 \; t_3 \; \dots \; t_N \right]$ is called the \textbf{target vector}.

We are, then, left with the following question: how can the model quantitatively evaluate the quality of its current set of parameters? That could be achieved in a number of ways, but the most usual is using a \textbf{Cost Function} that, as the name suggests, measures the cost of each wrong classification of the model. The model then evolves in a way that minimizes the cost function. A usual choice for the cost function is the \textbf{sum of squares error}. Mathematically, if $y_{\theta}^i = y_{\theta}(x_i)$ is the prediction for the input data point $x_i$ with label $t_i$, given the current set of parameters $\theta$, the cost function using this metric is given by

@@

Artificial Neural Networks (ANN) are mathematical structures that, as the name suggests, try to mimic the Human Brain. ANN's building blocks, like their biological counterpart, model the high-level behaviour of biological neurons, in the sense that they neglect unnecessary biological aspects (such as modeling all the voltages across the neuron and all its electromagnetic interactions), and only retain its fundamental underlying mathematical function, which is a weighted linear combination of its inputs subject to a \emph{activation function}, i.e. a function that outputs a decision value depending on its inputs.  Mathematically, we have

\begin{equation}
	y = f(\mb{w}^T \mb{x} + b_0)
\end{equation}
where $\mb{w} = \left[ w_1 \; w_2 \;  w_3 \;  \dots \; w_n \right]$ is the input weight vector, $\mb{x} = \left[ x_1 \; x_2 \; x_3 \; \dots \; x_n \right]$ the input vector, $b_0$ is the bias factor and $f(\cdot)$ is the chosen activation function. Furthermore, we call the scalar quantity $a = \mb{w}^T \mb{x} + b_0$ the \textbf{activation}, since its value determines how the activation function will behave. Figure~\ref{fig:modelNeuron} exemplifies the roles of these variables within our neuron model, and compares each part of it with the biological counterpart.

@@

    \caption{In Figure~\ref{fig:modelNeuron_a}. each input feature $x_i$ is weighted by its corresponding weight $w_i$. During the training procedure, these weights are adjusted so that the output $y$ approaches the target value. In Figure~\ref{fig:modelNeuron_b}, we see the diagram of an actual multi polar neuron. The dendrites, where the stimuli are received, plays a role similar to that of the input nodes. The axon transmits the signal to the synaptic terminals, that are similar to the $y$ output. Source: \href{https://en.wikipedia.org/wiki/Neuron\#/media/File:Neuron\_Hand-tuned.sv://en.wikipedia.org/wiki/Neuron\#/media/File:Blausen\_0657\_MultipolarNeuron.png}{Wikipedia}}
	\label{fig:modelNeuron}
\end{figure}

@@

\begin{figure}[H]
	\centering
	\includegraphics[width=0.9\linewidth]{figures/activFunc.eps}
	\caption{Three different activation functions. As you can see, the hyperbolic tangent has the same extreme value as the sign step function, but has a smooth transition between them, which can be interpreted as a \emph{soft decision} in the more ambiguous middle region, reflecting the degree of uncertainty on the decision. On the other hand, the sigmoid function goes from zero to one, and is also smooth like the hyperbolic tangent}.
	\label{fig:activFunc}
\end{figure}

@@

\begin{figure}[H]
	\centering
	\includegraphics[width=0.9\linewidth]{figures/ann.eps}
	\caption{A three layer ANN. We have omitted some of the connections in the hidden layer, for simplification purposes. $\mb{w}_1$ represents the weight matrix of the input layer, $\mb{w}_2$ the weight matrix of the connections between the input layer and the hidden layer, and $\mb{w}_3$ the weight connections between the hidden and the output layer. $f_{ij}(\dots)$ is the activation function of the $j$-th neuron of the $i$-th layer. Since they can be different, I chose different indexes to each.}
	\label{fig:neuralnet}
\end{figure}

Regarding the training of ANNs, is is performed through a two-step process: first, a \textbf{feed-forward} step where the input is applied, and the activations are evaluated in succession up to the output neurons; then, the perform the \textbf{backpropagation} step, where we calculate the errors in each of the nodes, but now from the output to the input: the weights are updated and optimized using an iterative  method called \textit{Gradient Descent}, where if $\tau$ is the current time step, the next update on the weight matrix $\mb{w}^{(\tau+1)}$ is given by

@@

\subsection{Recurrent Neural Networks}\label{sec:theorBack_rnn}
A Recurrent Neural Network (RNN) is, essentially, a regular ANN where some neurons (especially in the hidden layer) have \emph{feedback connections} to themselves, i.e. their outputs are fed as inputs. The relevance of this different structure is the possibility to retain \emph{sequence} information about the data. Before, each incoming data point only contributed to the training of the network, but no information about the correlation between themselves and the data points that preceded them did not influence the training step. They were regarded as if no temporal relationship existed, and therefore each data point is conditionally independent of any other. This is obviously not necessarily truth, and in fact there are many cases, where the correlation between data points is high for those closely spaced in time, in which it is actually completely false, as in video signals, audio signals, or other kinds of \emph{temporal sequences} of data. Therefore, the feedback connection of the neuron to himself acts as a kind of \emph{memory element} that takes into account in the present decision, the history of decisions previously taken, and hence the previous data. 
Figure~\ref{fig:recneuron} suggests a possible structure for a neuron of a hidden layer in an RNN, and also an alternate representation, where the structure is unfold through time.

%% METE AQUI A FIGURA

The training of an RNN is usually performed using a variation of Backpropagation, called \textbf{Backpropagation Through Time} (BPTT), that as the name implies, performs the same backpropagation procedure discussed for the ANNs, but now taking into account the unfolding of the network through a fixed training epoch $T$ like Figure~\ref{fig:recneuron}. Due to this very fact, this training procedure is memory and performance consuming, and so it will not be used in my final work, but instead a novel approach, the \textbf{Simulatneous Perturbation Stochastic Approximation}.

@@

\subsection{Long Short-Term Memory Networks}\label{sec:theorBack_lstm}
To overcome the issue of failing to remember long-term dependencies, Hochreiter and Schmidhüber proposed, in 1997, a novel approach to the RNNs called the \textit{Long-Short Term Network}~\cite{Hoch97}. This section explains the main idea of this approach~\ref{sec:struct_lstm}, and also how it is trained\ref{sec:training_lstm}, serving as a support basis for the work of this thesis.
Although originally formulated in 1997, its formulation has been incrementally updated in~\cite{Gers00} and~\cite{Gers2000}, and the most current version is the one in~\cite{Graves05}. One of the inital proposers of LSTM, Prof. Jürgen Schmidhüber, did a survey on the most common variations of the structure~\cite{Greff15} last year, and this will the basis of this short theoretical presenation, as well as the work that will be developed in this thesis.

@@

\subsubsection{Training -- SPSA}\label{sec:training_lstm}
Since this work aims to perform on-chip learning, it is important to find a suitable learning scheme. Since I am aiming at an hardware implementation, and although the memory resources of nowadays FPGAs are abundant, one must find an algorithm that uses the less memory as possible (at the smallest performance cost possible), in order to use the additional memory, for instance, to add additional LSTM cells that can improve the performance of our system. 
That being said, we see that the usual training algorithms for LSTM cells -- i.e. Real-Time Recurrent Learning (RTRL), Truncated Backpropagation Through Time (BPTT) or a mixture of both~\cite{Graves05}~\cite{Hoch97}~\cite{Greff15} -- usually involve the storage of the deltas of each layer for every time instant in the training epoch (from $0$ to $T$), which is a highly non-scalable solution both in terms of memory and performance. A most efficient approach to training times series dependant structures like LSTM is the use of \textbf{Simultaneous Perturbation Stochastic Approximation} (SPSA)~\cite{Spall98}.  The main idea of this technique is, instead of explicitly evalute the gradients for the cost function at each time step, perform a random walk in the neighbourhood of the current weight matrix, in the weight space, and approximate the new weight update by the approximation of the gradient of the cost function that resulted from that random walk. The weight update for the $i$-th weight at the time step $t$ is given by

@@

One last comment to be made concerning the update rule, is the hipothetical need for \textbf{limits} to the weight values, when the update rule exceeds $|w_{\text{max}}|$ (in that case, we set $w_i^{(t+1)} = \pm w_{\text{max}}$), which sometimes might be needed if the behaviour of the weight update is not appropriate~\cite{Maeda05}.

\section{Proposed Solution}\label{sec:propSol}
Given the previous discusson of how LSTM neurons work, their benefits when compared to other RNN architectures, this work aims to conceive an efficient dedicated FPGA implementation with on-chip learning that, hopefully, surpasses the training/testing time of an equivalent CPU implementation in a general high-level language or from a framework.

@@

This solution will combine the benefits of a dedicated hardware implementation, where the inherent parallelism so obviously visible in the network architecture can be fully exploited something that, predominantly, yields a much higher performance in terms of computation time than a naïve CPU implementation, as in~\cite{Maeda05} where a $21\times$ speed-up against an ARM CPU was acheived.

>>>>

Let us consider a practical example. For instance, suppose we want to build a program that given an input audio waveform representation of a spoken word, it matches it into a particular word of a dictionary. We could, of course, devise a set of rules and exceptions for each word analysing some of its features (perhaps the Fourier representation of each one, and, from it, manually finding the appropriate rules for each), but apart from it being a very complex task, it wouldn't be a scalable solution, given the enormous number of words in each language. The approach taken by Machine Learning is different, and instead of manually processing each waveform, we build a large dataset, of size $N$, containing the waveforms of several words $\left[ \mb{x}_1 \; \mb{x}_2 \;\mb{x}_2 \; \dots \mb{x}_N \right]$ -- we call this dataset the \textbf{training dataset} -- and we feed it to our model. Each of the $i$-th data point was previously labelled, and in fact we feed each training data point $\mb{x}_i$ \emph{along} with its corresponding label $t_i$, so that the model can adapt its parameters accordingly to the \emph{target value} it is supposed to classify. This set of labels $\mb{t} = \left[ t_1 \; t_2 \; t_3 \; \dots \; t_N \right]$ is called the \textbf{target vector}.

We are, then, left with the following question: how can the model quantitatively evaluate the quality of its current set of parameters? That could be achieved in a number of ways, but the most usual is using a \textbf{Cost Function} that, as the name suggests, measures the cost of each wrong classification of the model. The model then evolves in a way that minimizes the cost function. A usual choice for the cost function is the \textbf{sum of squares error}, given the Gaussian Noise assumption. Mathematically, if $y_{\theta}^i = y_{\theta}(x_i)$ is the prediction for the input data point $x_i$ with label $t_i$, given the current set of parameters $\theta$, the cost function using this metric is given by

@@

Artificial Neural Networks (ANN) are mathematical structures that, as the name suggests, try to mimic the basic way of how a human brain works. ANN's building blocks, like their biological counterpart, model the high-level behaviour of biological neurons, in the sense that they neglect unnecessary biological aspects (such as modeling all the voltages across the neuron and all its electromagnetic interactions), and only retain its fundamental underlying mathematical function, which is a weighted linear combination of its inputs subject to a \emph{activation function}, i.e. a function that outputs a decision value depending on its inputs.  Mathematically, we have

\begin{equation}
	y = f(\mb{w}^T \mb{x})
\end{equation}
where $\mb{w} = \left[ w_0 \; w_1 \; w_2 \;  w_3 \;  \dots \; w_n \right]$ is the input weight vector, $\mb{x} = \left[ x_0 \; x_1 \; x_2 \; x_3 \; \dots \; x_n \right]$ the input vector, $b_0$ is the bias factor and $f(\cdot)$ is the chosen activation function. Furthermore, we call the scalar quantity $a = \mb{w}^T \mb{x}$ the \textbf{activation}, since its value determines how the activation function will behave. Figure~\ref{fig:modelNeuron} exemplifies the roles of these variables within our neuron model, and compares each part of it with the biological counterpart.

@@

    \caption[Representations of a Neuron in an ANN~\ref{fig:modelNeuron_a} and a biological multi polar neuron~\ref{fig:modelNeuron_b}]{In Figure~\ref{fig:modelNeuron_a}. each input feature $x_i$ is weighted by its corresponding weight $w_i$. During the training procedure, these weights are adjusted so that the output $y$ approaches the target value. In Figure~\ref{fig:modelNeuron_b}, we see the diagram of an actual multi polar neuron. The dendrites, where the stimuli are received, plays a role similar to that of the input nodes. The axon transmits the signal to the synaptic terminals, that are similar to the $y$ output. Source: \href{https://en.wikipedia.org/wiki/Neuron\#/media/File:Neuron\_Hand-tuned.sv://en.wikipedia.org/wiki/Neuron\#/media/File:Blausen\_0657\_MultipolarNeuron.png}{Wikipedia}}
	\label{fig:modelNeuron}
\end{figure}

@@

\begin{figure}[H]
	\centering
	\includegraphics[width=0.9\linewidth]{figures/activFunc.eps}
    \caption[Possible activation functions of choice for a neuron]{Three different activation functions. As you can see, the hyperbolic tangent has the same extreme value as the sign step function, but has a smooth transition between them, which can be interpreted as a \emph{soft decision} in the more ambiguous middle region, reflecting the degree of uncertainty on the decision. On the other hand, the sigmoid function goes from zero to one, and is also smooth like the hyperbolic tangent}.
	\label{fig:activFunc}
\end{figure}

@@

\begin{figure}[H]
	\centering
	\includegraphics[width=0.9\linewidth]{figures/ann.eps}
    \caption[Three layer ANN]{A three layer ANN. We have omitted some of the connections in the hidden layer, for simplification purposes. $\mb{w}_1$ represents the weight matrix of the input layer, $\mb{w}_2$ the weight matrix of the connections between the input layer and the hidden layer, and $\mb{w}_3$ the weight connections between the hidden and the output layer. $f_{ij}(\dots)$ is the activation function of the $j$-th neuron of the $i$-th layer. Since they can be different, I chose different indexes to each.}
	\label{fig:neuralnet}
\end{figure}

Regarding the training of ANNs, it is performed through a two-step process: first, a \textbf{feed-forward} step where the input is applied, and the activations are evaluated in succession up to the output neurons; then, we perform the \textbf{backpropagation} step, where we calculate the errors in each of the nodes (the so-called \textbf{deltas} of equation~\ref{eq:deltaj}), but now from the output to the input: the weights are updated and optimized using an iterative  method called \textit{Gradient Descent}, where if $\tau$ is the current time step, the next update on the weight matrix $\mb{w}^{(\tau+1)}$ is given by

@@

\subsection{Recurrent Neural Networks}\label{sec:theorBack_rnn}
A Recurrent Neural Network (RNN) is, essentially, a regular ANN where some neurons (especially in the hidden layer) have \emph{feedback connections} to themselves, i.e. their outputs are fed as inputs. The relevance of this different structure is the possibility to retain \emph{sequence} information about the data. Before, each incoming data point only contributed to the training of the network, but the information about the correlation between themselves and the data points that preceded them did not influence the training step. Temporal relationship is disregarded and each data point considered conditionally independent of any other. This is obviously not necessarily true, and in fact there are many cases where the correlation between data points is high for those closely spaced in time, such as in video signals, audio signals, or other kinds of \emph{temporal sequences} of data. Therefore, the feedback connection of the neuron to himself acts as a kind of \emph{memory element} that takes into account in the present decision, the history of decisions previously taken, and hence the previous data.

Figure~\ref{fig:recurrneuron} suggests a possible structure for a neuron of a hidden layer in an RNN, and also an alternate representation, where the structure is unfold through time.

\begin{figure}[H]
	\begin{subfigure}{0.5\linewidth}
		\centering
		\includegraphics[width=0.34\linewidth]{figures/recneuron.eps}
		\caption{Recurrent Neuron}
		\label{fig:recneuron}
	\end{subfigure}
	\begin{subfigure}{0.5\linewidth}
		\centering
		\includegraphics[width=0.9\linewidth]{figures/recneuron_unf.eps}
		\caption{Recurrent Neuron, unfolded through time}
		\label{fig:recneuron_unf}
	\end{subfigure}

    \caption[Representation of a Recurrent Neuron (Figure~\ref{fig:recneuron}) and a Recurrent Neuron unfolded through time (Figure~\ref{recneuron_unf})]{In Figure~\ref{fig:recneuron}, a recurrent neuron is depicted, where the output is fed back to the input, weighted by $\theta_h$. Figure~\ref{fig:recneuron_unf}, an unfolding through time of that same neuron is performed. The basic idea is that feeding the output back to the input is similar to feeding the output to the input of the neuron \emph{at a previous time step}: this way, we linearize the structure. Considering a training epoch of $T$ samples, corresponds to having $T$ unfolded neurons/layers}
	\label{fig:recurrneuron}
\end{figure}

The training of an RNN is usually performed using a variation of Backpropagation, called \textbf{Backpropagation Through Time} (BPTT), that as the name implies, performs the same backpropagation procedure discussed for the ANNs, but now taking into account the unfolding of the network through a fixed training epoch $T$ like Figure~\ref{fig:recneuron_unf}. Due to this very fact, this training procedure is memory and performance consuming, and so it will not be used in my final work, but instead a novel approach, the \textbf{Simultaneous Perturbation Stochastic Approximation}.

@@

\subsection{Long Short-Term Memory Networks}\label{sec:theorBack_lstm}
To overcome the issue of failing to remember long-term dependencies, Hochreiter and Schmidhüber proposed, in 1997, a novel approach to the RNNs called the \textit{Long-Short Term Network}~\cite{Hoch97}. This section explains the main idea of this approach~\ref{sec:struct_lstm}, and also how it is trained\ref{sec:training_lstm}, serving as a support for the work of this thesis.
Although originally formulated in 1997, its formulation has been incrementally updated in~\cite{Gers00} and~\cite{Gers2000}, and the most current version is the one in~\cite{Graves05}. One of the inital proposers of LSTM, Prof. Jürgen Schmidhüber, did a survey on the most common variations of the model last year~\cite{Greff15}, and this will the basis of this short theoretical presenation, as well as the work that will be developed in this thesis.

@@

\subsubsection{Training -- SPSA}\label{sec:training_lstm}
Since this work aims to perform on-chip learning, it is important to find a suitable learning scheme. Since I am aiming at an hardware implementation, and although the memory resources of nowadays FPGAs are abundant, one must find an algorithm that uses the less memory as possible (at the smallest performance cost possible), in order to use the additional memory, for instance, to add additional LSTM cells that can improve the performance of our system. 
That being said, we see that the usual training algorithms for LSTM cells -- i.e. Real-Time Recurrent Learning (RTRL), Truncated Backpropagation Through Time (BPTT) or a mixture of both~\cite{Graves05}~\cite{Hoch97}~\cite{Greff15} -- usually involve the storage of the deltas of each layer for every time instant in the training epoch (from $0$ to $T$), which is a highly non-scalable solution both in terms of memory and performance. A most efficient approach to training times series dependant structures like LSTM is the use of \textbf{Simultaneous Perturbation Stochastic Approximation} (SPSA)~\cite{Spall98}.  The main idea of this technique is, instead of explicitly evaluate the gradients for the cost function at each time step, perform a random walk in the neighbourhood of the current weight matrix, in the weight space, and approximate the new weight update by the approximation of the gradient of the cost function that resulted from that random walk. The weight update for the $i$-th weight at the time step $t$ is given by

@@

One last comment to be made concerning the update rule, is the hypothetical need for \textbf{limits} to the weight values, when the update rule exceeds $|w_{\text{max}}|$ (in that case, we set $w_i^{(t+1)} = \pm w_{\text{max}}$), which sometimes might be needed if the behaviour of the weight update is not appropriate~\cite{Maeda05}.

\section{Proposed Solution}\label{sec:propSol}
Given the previous discussion of how LSTM neurons work and their benefits when compared to other RNN architectures, this work aims to conceive an efficient dedicated FPGA implementation with on-chip learning that, hopefully, surpasses the training/testing time of an equivalent CPU implementation in a general high-level language or of a framework.

@@

This solution will combine the benefits of a dedicated hardware implementation, where the inherent parallelism so obviously visible in the network architecture can be fully exploited something that, predominantly, yields a much higher performance in terms of computation time than a naïve CPU implementation (as in~\cite{Maeda05} where a $21\times$ speed-up against an ARM CPU was acheived) with the flexibility that a reconfigurable platform such as an FPGA provides.
